---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip(zipfile='activity.zip')
activity <- read.csv('activity.csv')
```

Summary of the csv file:

```{r}
summary(activity)
```

## What is mean total number of steps taken per day?

Total number of steps taken per day:
```{r}
aggregated_steps <- aggregate(steps ~ date, activity, sum, rm.na = TRUE)
```

plot a histogram:
```{r}
library(ggplot2)
g <- ggplot(aggregated_steps,aes(x=steps))+geom_histogram(stat="bin",bins=20)
print(g)
```

Mean and median of the number of steps:
```{r}
mean(aggregated_steps$steps,na.rm=TRUE)

median(aggregated_steps$steps,na.rm=TRUE)
```


## What is the average daily activity pattern?

Time series:
```{r}
average_steps <- aggregate(steps ~ interval, activity, mean, rm.na = TRUE)
g <- ggplot(average_steps, aes(x=interval,y=steps))+geom_line()
print(g)
```

Maximum number of average steps:
```{r}
average_steps[order(average_steps$steps,decreasing = TRUE),][1,]

```


## Inputing missing values

Number of rows that contain NA:
```{r}
nrow(activity[is.na(activity),])
```
 
Compensate NA values by the mean of that interval:
```{r}
intervals <-  unique(activity[,3])
for (i in 1:length(intervals)){
   v <- average_steps[average_steps$interval == intervals[i],]$steps
   activity[is.na(activity$steps) & activity$interval == intervals[i],]$steps <- v
  
}
``` 

New histogram:
```{r}
aggregated_steps <- aggregate(steps ~ date, activity, sum, rm.na = TRUE)
g <- ggplot(aggregated_steps,aes(x=steps))+geom_histogram(stat="bin",bins=20)
print(g) 
mean(aggregated_steps$steps,na.rm=TRUE)
median(aggregated_steps$steps,na.rm=TRUE)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday_factor <- function(date) {
    if (weekdays(as.Date(date,'%Y-%m-%d')) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
activity$weekday <- sapply(activity$date,weekday_factor)
sum_steps_weekdays <- aggregate(steps ~ interval + weekday, activity, sum, rm.na = TRUE)
g <- ggplot(sum_steps_weekdays, aes(x=interval,y=steps,color=weekday))+geom_line()+facet_wrap(~weekday)
print(g)
```
